---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- badges: start -->
  [![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
  <!-- badges: end -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(tidyverse.quiet = TRUE)
```

# gglobalclocks

What if you could draw a beautiful 'wall' of clocks with local times from around the world?  Well, gglobalclocks lets you do that!

I worked on the wall of clocks thinking it might help me schedule virtual meetings with participant in lots of different time zones.  Does a beautiful wall of clocks help schedule global meetings?  As it turns out, in my experience, not really!

But gglobalclocks also has utilities for building dataframes of locations and local times which I have found helpful for multi-timezone scheduling.


```{r, eval = T}
library(tidyverse)
gglobalclocks:::date_time_tz_to_tzs() |> 
  gglobalclocks:::local_tzs_df_collapse() |>
  gglobalclocks:::gglobalclocks() + 
    aes(local_time = local_time) + 
    gglobalclocks:::stamp_clockface() + 
    gglobalclocks:::geom_clock_hands() + 
    facet_wrap(~str_wrap(locations, 20))
  
tz_targets <- c("US/Mountain","US/Eastern","US/Eastern", "Europe/Paris", "Europe/Amsterdam", "America/Sao_Paulo", "America/Santiago", "Australia/Melbourne")

gglobalclocks:::date_time_tz_to_tzs(
  from_date_time = "2024-03-06 12:00:00",
  from_tz = "US/Mountain",
  to_tz = tz_targets)

df <- gglobalclocks:::date_time_tz_to_tzs(
  from_date_time = "2024-03-06 12:00:00",
  from_tz = "US/Mountain",
  to_tz = tz_targets) |> 
  gglobalclocks:::local_tzs_df_collapse()

df

df |>  
  ggplot() + 
  gglobalclocks:::stamp_workday() +
  aes(local_time, fct_inorder(locations)) + 
  geom_point() +
  geom_text(aes(label = local_time)) + 
  aes(color = local_wday_date)
```


```{r, comment=""}
library(tidyverse)
library(lubridate)
```

---

## Part II. Functions discussion and definitions

### `time_to_local()`

One helper function to do translation from one time zone to multiple time zones is time_to_local.  From working on global clocks, it seems like you can't keep date-times with different time zones in one vector (or variable), so you need to do a conversion one by one and save a character version of the complete time zone information.

```{r time_to_local}
time_to_local <- function(x, tz){
  lubridate::with_tz(x, tz = tz) %>%
    as.character()
}
```

### `date_time_tz_to_tzs()`

We can use the converter above to translate from a 'from' location and time to a bunch of locations' local times. We do this and then add a few more helpful columns like local_time, local_date etc.  

```{r date_time_tz_to_tzs}
date_time_tz_to_tzs <- function(from_date_time = "2024-03-27 12:00:00", 
                                from_tz = "US/Mountain", 
                                to_tz = c("Europe/Paris", 
                                          "Europe/Amsterdam", 
                                          "Australia/Adelaide",
                                          "Australia/Melbourne", 
                                          "Europe/Stockholm", 
                                          "US/Eastern",
                                          "Europe/London", 
                                          "US/Pacific",
                                          "US/Mountain", 
                                          "US/Central", 
                                          "Australia/Sydney",
                                          "Europe/Vienna",
                                          "America/Sao_Paulo",
                                          "America/Santiago",
                                          "America/Buenos_Aires")){

   meeting <- ymd_hms(from_date_time, 
                      tz = from_tz)

OlsonNames() %>%
  data.frame(tz = .) %>%
  dplyr::filter(tz != "US/Pacific-New") %>%
  dplyr::filter(tz %in% to_tz) %>%
  dplyr::mutate(local_date_time_chr = purrr::map2(meeting, tz, time_to_local))  %>%
  tidyr::unnest(local_date_time_chr) %>%
  dplyr::mutate(local_time_date_utc = 
                  lubridate::ymd_hms(local_date_time_chr, tz = "UTC"))  %>% 
  dplyr::mutate(local_time = lubridate::ymd_hms(local_date_time_chr))  %>% 
  dplyr::mutate(local_time = hms::as_hms(local_time)) %>%
  dplyr::mutate(local_date = as.Date(local_date_time_chr)) %>%
  dplyr::mutate(local_wday = lubridate::wday(local_date, label = T)) %>%
  dplyr::arrange(local_date, local_time) %>%
  dplyr::select(-local_date_time_chr) %>%
  dplyr::mutate(local_wday_date = paste0(local_wday, ", ", month(local_date, label = T), " ", day(local_date)))

}
```


```{r}
date_time_tz_to_tzs() |>
  head()
```

### `local_tzs_df_collapse()`

If it turns out that you have multiple tz with the same local meeting time, you can collapse these locations by local time.  

```{r local_tzs_df_collapse}
local_tzs_df_collapse <- function(local_tzs_df, collapse = "; "){
  
  local_tzs_df |>
    group_by(local_date, local_time, local_wday_date) |>
    summarise(locations = paste(tz, collapse = collapse)) |>
    ungroup() |>
    select(locations, everything()) |>
    mutate(locations = str_remove_all(locations, "Europe/|America/|Australia/")) |>
    arrange(local_date, local_time) |>
    mutate(location = fct_inorder(locations)) |>
    select(-local_date)
  
}
```

This is the type of information you might send to an attendees list so they can know, at a glance, their likelihood of making attending work.

```{r}
date_time_tz_to_tzs() |> 
  local_tzs_df_collapse() |> 
  head()
```

# More charming display...

You can of course display in a nicer way, passing to html table maker for example.

```{r}
date_time_tz_to_tzs() |> 
  local_tzs_df_collapse() |> 
  knitr::kable()
```


### Let's build a wall of global clocks with base ggplot2

```{r}
date_time_tz_to_tzs() |> 
  local_tzs_df_collapse() |>
  mutate(minute_turn = local_time %>% 
           lubridate::minute() %% 60 / 60) %>% 
  mutate(hour_turn = local_time %>% 
           lubridate::hour() %% 12/12 + minute_turn/12) %>% 
  mutate(am_pm = ifelse(local_time %>% lubridate::hour() >= 12, "pm", "am")) %>% 
  ggplot() + 
  facet_wrap(~ str_wrap(locations, 15)) + 
  aes(x =  hour_turn, 
      xend =  hour_turn,
      y = 0,
      yend = .6,
      color = am_pm) + 
  labs(color = NULL) +
  geom_segment() + 
  geom_segment(aes(x =  minute_turn, 
                  xend =  minute_turn ,
               yend = 1) ) + 
  coord_polar() + 
  scale_x_continuous(limits = c(0,1)) + 
  geom_text(data = tibble(x = 1:12, y = 1), 
            aes(label = x, x = x/12, y = y, xend = NULL, 
                yend = NULL, color = NULL),
            show.legend = F) + 
  theme_void() + 
  annotate(geom = "segment", x = 0, xend = 1, y = 1.2, yend = 1.2)
```

## Speed things up by putting some of this into functions

### `gglobalclocks()`

```{r gglobalclocks, eval = T}
gglobalclocks <- function(data = NULL){
  
  ggplot2::ggplot(data) + 
  list(ggplot2::theme_void(),
       ggplot2::coord_polar(),
       ggplot2::scale_x_continuous(limits = c(0,1)),
       ggplot2::scale_y_continuous(limits = c(0,1.3)))
  
}
```

### `stamp_clockface()`


```{r stamp_clockface}
stamp_clockface <- function(){
  
    list(ggplot2::geom_text(data = data.frame(x = 1:12, y = 1), 
            ggplot2::aes(label = x, x = x/12, y = y, xend = NULL, 
                yend = NULL, color = NULL, local_time = NULL),
            show.legend = F),
          ggplot2::annotate(geom = "segment", x = 0, xend = 1, y = 1.2, yend = 1.2)
         
    )
  
}
```

```{r}
gglobalclocks() + stamp_clockface()
```

### `statClockhourhand`


```{r statClockhourhand}
# there's probably a lot of fun to be had refactoring this!
# seems quite verbose
compute_clock_hands <- function(data){
  
  data |> 
  dplyr::mutate(minute_turn = local_time %>% lubridate::minute() %% 60 / 60) |> 
  dplyr::mutate(hour_turn = local_time %>% lubridate::hour() %% 12/12 + minute_turn/12) |>
  dplyr::mutate(am_pm = ifelse(local_time %>% lubridate::hour() > 12, "pm", "am")) 
  
}

compute_hour_hand <- function(data, scales){
  
  data |>
    compute_clock_hands() |>
    dplyr::mutate(x =  hour_turn, 
      xend =  hour_turn,
      y = 0,
      yend = .6)
  
}

compute_minute_hand <- function(data, scales){
  
  data |>
    compute_clock_hands() |>
    dplyr::mutate(x =  minute_turn, 
      xend =  minute_turn,
      # color = am_pm, 
      y = 0,
      yend = 1)
  
}


statClockminhand <- ggplot2::ggproto(`_class` = "statClockminhand",
                          `_inherit` = ggplot2::Stat,
                          # required_aes = c("local_time"),
                          compute_group = compute_minute_hand,
                          default_aes = ggplot2::aes(color =
                                                            ggplot2::after_stat(am_pm))
                           )

geom_minute_hand <- function(
  mapping = NULL,
  data = NULL,
  position = "identity",
  na.rm = FALSE,
  show.legend = NA,
  inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = statClockminhand,  # proto object from step 2
    geom = ggplot2::GeomSegment,  # inherit other behavior
    data = data,
    mapping = mapping,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}

statClockhourhand <- ggplot2::ggproto(`_class` = "statClockhourhand",
                          `_inherit` = ggplot2::Stat,
                          # required_aes = c("local_time"),
                          compute_group = compute_hour_hand,
                          default_aes = ggplot2::aes(color =
                                                            ggplot2::after_stat(am_pm))
                           )
```

```{r}
date_time_tz_to_tzs() |> compute_clock_hands()
date_time_tz_to_tzs() |> compute_minute_hand()
```

### `geom_clock_hands()`

```{r geom_clock_hands}
geom_hour_hand <- function(
  mapping = NULL,
  data = NULL,
  position = "identity",
  na.rm = FALSE,
  show.legend = NA,
  inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = statClockhourhand,  # proto object from step 2
    geom = ggplot2::GeomSegment,  # inherit other behavior
    data = data,
    mapping = mapping,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}


geom_clock_hands <- function(...){
  
  list(geom_hour_hand(...),
       geom_minute_hand(...))
  
}
```

### Try it out

```{r}
date_time_tz_to_tzs() |>
  local_tzs_df_collapse() |>
gglobalclocks() +
  aes(local_time = local_time) +
  stamp_clockface() +
  geom_minute_hand() +
  geom_hour_hand() +
  facet_wrap(~locations)

date_time_tz_to_tzs() |> 
  local_tzs_df_collapse() |>
  gglobalclocks() + 
  aes(local_time = local_time) + 
  stamp_clockface() + 
  geom_clock_hands() + 
  facet_wrap(~locations)
```

```{r}
readme2pkg::chunk_to_r("gglobalclocks")
readme2pkg::chunk_to_r("stamp_clockface")

readme2pkg::chunk_to_r("statClockhourhand")
readme2pkg::chunk_to_r("geom_clock_hands")
```


## If a wall of clocks is unhelpful in scheduling, can we viz in a different way? Towards LocationXLocaTime plot

### `stamp_workday()`

We'll use an 'stamp' (annotation) layer that helps us get oriented to the 12 hours in a work day.

```{r stamp_workday}
create_day_schedule_df <- function(){
  
  data.frame(time_start = 
               hms::as_hms(c("00:00:00","07:00:00", "09:00:00", "17:00:00","21:00:00")), 
             time_end = 
               hms::as_hms(c("07:00:00","09:00:00", "17:00:00", "21:00:00","24:00:00")),
             stance = c("avoid","awake", "business","awake","avoid"))
}


stamp_workday <- function(show.legend = T){  
  
  ggplot2::geom_rect(data = create_day_schedule_df(),
            ggplot2::aes(xmin = time_start,
                xmax = time_end, 
                x = NULL,
                fill = stance,
                ymin = -Inf,
                ymax = Inf,
                y = NULL,
                color = NULL),
           alpha = .5, show.legend = show.legend) 
}
```

```{r}
readme2pkg::chunk_to_r(chunk_name = "stamp_workday")
```

## LocationXLocalTime

```{r}
date_time_tz_to_tzs() |> 
  local_tzs_df_collapse() |>  
  ggplot() + 
  aes(local_time, fct_rev(fct_inorder(str_wrap(locations,25)))) + 
  labs(x = "Local meet time", y = NULL) + 
  stamp_workday() +
  geom_point() + 
  geom_text(aes(label = local_time),
            hjust = -.1, show.legend = F) + 
  geom_vline(xintercept = hms::as_hms("12:00:00"),
              linetype = "dashed", color = "grey25",
              alpha = .2) +
  aes(color = local_wday_date) +
  labs(color = "Local meet date") +
  theme(legend.position = "top", 
        legend.justification = "left") + 
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_line(color = ""))


```



# Part II. Packaging and documentation  🚧

## Phase 1. Minimal working package

### Bit A. Created package archetecture, running `devtools::create(".")` in interactive session. ✅


```{r, eval = F}
devtools::create(".")
```

### Bit B. Added roxygen skeleton? 🚧 

Use a roxygen skeleton for auto documentation and making sure proposed functions are *exported*.  Generally, early on, I don't do much (anything) in terms of filling in the skeleton for documentation, because things may change.

### Bit C. Managed dependencies ? 🚧 

Package dependencies managed, i.e. `depend::function()` in proposed functions and declared in the DESCRIPTION

```{r, eval = F}
usethis::use_package("dplyr")
usethis::use_package( "hms")
usethis::use_package( "lubridate")
usethis::use_package( "purrr")
usethis::use_package( "tidyr")
```


### Bit D. Moved functions R folder? ✅

Use new {readme2pkg} function to do this from readme...

```{r}
readme2pkg::chunk_to_r("date_time_tz_to_tzs")
readme2pkg::chunk_to_r("local_tzs_df_collapse")
readme2pkg::chunk_to_r("time_to_local")
```

### Bit E. Run `devtools::check()` and addressed errors. 🚧 

```{r, eval = F}
devtools::check(pkg = ".")
```

### Bit F. Build package 🚧 ✅

```{r, eval = F}
devtools::build()
```


### Bit G. Write traditional README that uses built package (also serves as a test of build. ✅ 

The goal of the {ggchalkboard} package is to make it easy to theme ggplots like chalkboards

Install package with:

```
remotes::install_github("EvaMaeRey/ggchalkboard")
```

Once functions are exported you can remove go to two colons, and when things are are really finalized, then go without colons (and rearrange your readme...)

```{r, eval = F}
library(gglobalclocks)  

gglobalclocks:::date_time_tz_to_tzs("2024-03-06 11:00:00", 
                                    from_tz = "US/Mountain",
                                    to_tz = c("US/Eastern",
                                              "US/Mountain",
                                              "US/Pacific",
                                              "Europe/Paris"))
```




### Bit H. Chosen a license? ✅


```{r, eval = F}
usethis::use_mit_license()
```

### Bit I. Add lifecycle badge (experimental) ✅

```{r, eval = F}
usethis::use_lifecycle_badge("experimental")
```


## Phase 2: Listen & iterate 🚧 

Try to get feedback from experts on API, implementation, default decisions.  Is there already work that solves this problem? 


## Phase 3: Let things settle

### Bit A. Settle on examples.  Put them in the roxygen skeleton and readme. 🚧 

### Bit B. Written formal tests of functions and save to test that folders 🚧 


That would look like this...

```{r test_calc_times_two_works, eval = F}
library(testthat)

test_that("calc times 2 works", {
  expect_equal(times_two(4), 8)
  expect_equal(times_two(5), 10)
  
})
```


```{r, eval = F}
readme2pkg::chunk_to_tests_testthat("test_calc_times_two_works")
```


### Bit C. Added a description and author information in the DESCRIPTION file 🚧 

### Bit D. Addressed *all* notes, warnings and errors. 🚧 

## Phase 4. Promote to wider audience...

### Bit A. Package website built?  ✅

```{r, eval = F}
usethis::use_pkgdown()
pkgdown::build_site()
```


### Bit B. Package website deployed? 🚧 ✅

## Phase 5: Harden/commit

### Submit to CRAN/RUniverse? 🚧 

# Appendix: Reports, Environment

## Edit Description file

```{r, eval = T}
readLines("DESCRIPTION")
```

## Environment

Here I just want to print the packages and the versions

```{r}
all <- sessionInfo() |> print() |> capture.output()
all[11:20]
```

## `devtools::check()` report

```{r, error = T, results="hide", warning=F, eval = F}
devtools::check(pkg = ".")
```

What would gglobalclocks syntax look like using above approach as jumping off point?



